Controlling the dewetting morphologies of thin liquid films by switchable substrates

### 2022-10-03 Reply_RefereeA.tex

\documentclass[12pt,english]{article}

\usepackage{natbib}

\usepackage{graphics,graphicx,dcolumn,bm,fleqn,epic,eepic,float}
\usepackage{amssymb,amsmath,multirow,rotate,rotating,color}
\usepackage[utf8]{inputenc}

\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\tikzset{fontscale/.style = {font=\relsize{#1}}}
\usetikzlibrary{calc}
\makeatother

\newcommand{\figref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\eqnref}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\ts}{\textsuperscript}

\definecolor{tuered}{RGB}{214,0,74}

\newcommand{\todo}[1]{\textbf{\textcolor{tuered}{ TODO: #1}}}

\newcommand{\vectornorm}[1]{\left|\left|#1\right|\right|}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\uvec}[1]{\hat{\vec{#1}}}
\newcommand{\tensor}[1]{\mathbf{#1}}

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}

\newcommand{\JH}[1]{\textcolor{blue}{JH: #1}}
\begin{document}

% The referee PDF has been generated from r239 -- 'make referee' command.

\section*{Reply to Referee A}
We thank the Referee for his/her report on our manuscript. 
We are glad to read that he/she finds it of potential interest and in principle 
suitable of publication, if a thorough revision is performed. 
We substantially improved and revised the paper for the resubmission and hope that we were able to clarify the obscure points, following all the Referee's remarks.
Hereafter we provide answers to the various questions/comments raised. All changes in the revised version are reported in red colour.

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{ 
I am unable to follow most of the arguments made with respect to the disjoining pressure. 
For me, one needs to integrate the disjoining pressure to arrive at an effective interface potential, and 
I do not really know how to go from the effective interface potential to the surface tension. 
The equation (3) therefore merits much more discussion.
}

\item[ \textbf{{Answer}}]
{
We did not give, in fact, any particular argument with respect to the disjoining pressure, since 
we assumed that it is a very well established concept employed in any theoretical and computational study of thin liquid film hydrodynamics in the lubrication approximation framework.
We acknowledge, though, that our assumption was probably erroneous and we give some more explanation. 
First of all, there is no need (and no way) to go from the disjoining pressure (or the interface potential) to the surface tension, since the two quantities carry complementary physical information. 
Let us, first, make a step back and underline a fundamental difference between the physics of {\it one free interface} and that of a {\it film}:
the latter is a layer of liquid confined between {\it two} interfaces. When the two interfaces are brought close enough to each other (the layer is thin), they interact.
The disjoining pressure stems from this interaction. Thermodynamically, it represents minus the derivative of the free energy difference, per unit surface, between 
the film and the bulk of the same liquid, or, equivalently, the pressure difference inside and outside the film 
\cite{Deryaguin1940,DeryaguinChuraev1978}.
In the case of wetting, there can be three types of interfaces: liquid/air, liquid/solid (the substrate) and, 
if the substrate is only partially wetted, solid/air, thus giving rise to three-phase contact lines. 
The usually called {\it surface tension} is the energy cost of forming a liquid/air interface of unit area and its only knowledge is, therefore, not enough in our case. 
The Young's condition \cite{Young1805,DeGennes1985} connects the surface tension with the analogous energies per unit
area associated with the solid/liquid ($\gamma_{\text{sl}}$) and solid/gas ($\gamma_{\text{sg}}$) interface as
\begin{equation}
\cos \theta = \frac{\gamma_{\text{sg}}-\gamma_{\text{sl}}}{\gamma},
\end{equation}
where $\theta$ is the contact angle.
So, there is no need (and no way) to go from the disjoining pressure (or the interface potential) to the surface tension, since the two quantities carry complementary physical information,
and both are required to determine the dynamics of dewetting.
Of course, on the other hand, the disjoining pressure depends also on the surface tension, which appears, in fact, in its expression (3). 
We think, though, that part of the confusion might be due to the fact that we do not define the symbol of the surface tension at its first occurrence, but only later. 
We have now fixed this and we apologize for it.
\JH{Do we need to change more in the paper?}
}

\item[ \textbf{\underline{Comment 2.}}] 
{
The same holds for the time scale. 
The viscous time scale is well defined, but necessitates a length scale in addition to the 
capillary velocity. It is unclear why the authors choose the length scale that they use in the paper.
Subsequently, they replace the surface tension by a complicated partial derivative of the 
disjoining pressure; here the same remark as above holds.
}

\item[ \textbf{{Answer}}]
{
We apologize if we have not been thorough in the discussion of the relevant scales, which indeed deserves some clarification. 
Thin film hydrodynamics is an intrinsically multiscale phenomenon, whereby there are at least two characteristic length scales: the mean film thickness (let us call it $h_0$)
and an horizontal length scale $\ell$ (which in our case can be, for example, the pattern wavelength, $\lambda$), where, typically $h_0 \ll \ell$. We chose to use the mean 
film thickness as a scale, because it is the only relevant length that is common to all the simulations.
In the same spirit, we chose as a reference time $t_0(\theta)$, which is not exactly the capillary time (though being related to it). 
$t_0$, as we explain in the text, is, in fact, the characteristic time of growth of the most unstable mode in the spinodal dewetting on a homogeneous
substrate of contact angle $\theta$. In spite of the substrate being patterned in our case, it is possible to define an effective $t_0$ in terms 
of the highest contact angle on the pattern, which is what determines the fastest instability, i.e. $t_0(\Theta)$, where 
$\Theta=\mbox{max}_{\mathbf{x}}\{\theta(\mathbf{x})\}= \theta_0 + \delta \theta$. We deem the so defined characteristic time as an appropriate scale,
since it does not depend on the pattern wavelength and speed. Actually, one can see that it captures well the time scales of the dewetting process
from figure 2, where it is shown that the rupture times are of the order of fraction to units of $t_0$.
We recognize, on the other hand that in this perspective the choice of $v_0 = \lambda/t_0$ as a reference velocity looks a bit awkward.
A more natural candidate would be, indeed, $v_0=h_0/t_0$. We have therefore decided to use the latter in the revised version of the manuscript.
We have also rephrased the corresponding paragraph in the {\it Method} section as follows: \\

\textcolor{red}{Hereafter, lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film. On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads 
$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}; in our patterned case, we define 
$q_0^2=h_{\ast}^{-2}(1-\cos\Theta)f^{\prime}(h_0/h_{\ast})$, where 
$\Theta = \max_{\mathbf{x}} \theta(\mathbf{x}) = \theta_0 + \delta \theta$.
A velocity scale also then naturally arises, namely $v_0 = h_0/t_0$.}

}

\item[ \textbf{\underline{Comment 3.}}]
{ 
Incidentally, for a 'normal' fluid such as water, the capillary velocity is $\sim 70\mbox{m}/\mbox{s}$, 
so that very, very rapidly the fluid inertia becomes important. 
This is indeed what one usually observes for the formation or coalescence of drops, so that at the very least this should be discussed and the Reynolds number calculated.
}

\item[ \textbf{{Answer}}]
{
Such a high value corresponds to the ratio of (surface tension)/(dynamic viscosity) for water, i.e. $v = \gamma/\eta \approx 70 \mbox{m}/\mbox{s}$.
This expression, though, is only representative of the capillary velocity, again, for {\it free surface flows}, as, e.g., in the coalescence of 
two droplets in the bulk of another fluid. 
In the case wetting/dewetting, 
the phenomenology is complicated by the liquid/substrate (and vapour/substrate) interaction. 
In such a situation the relevant velocity scale is set by the characteristic speed of contact line 
motion which can be estimated from a viscocapillary balance as 
$\sim \varepsilon^3 \gamma/\eta$, where $\varepsilon = h_0/L \ll 1$ is the film thickness parameter 
\cite{RevModPhys.69.931,doi:10.1146/annurev-fluid-011212-140734}. 
The typical velocities in dewetting problems are, then, much smaller than the ratio $\gamma/\eta$.
Moreover, in many actual applications, as well as in the majority of experiments, the 
working fluids are in general much more viscous than water (e.g., polymeric liquids), thus making
the characteristic velocity even smaller.
For the dewetting of a millimetric film of micrometric thickness reatracting into a single droplet, for instance, experiments have measured that 
the typical velocity is $\sim 1 \mbox{mm}/\mbox{s}$, which gives a Reynolds number $Re \approx 10^{-3}$ \cite{Edwards2016}. 
In our problem, there are two relevant characteristic velocities: the capillary velocity, $U_{\text{cap}}$, and 
the pattern speed, $v_{\theta}$. When their ratio $\Gamma = v_{\theta}/U_{\text{cap}} < 1$, the capillary 
velocity dominates and fixes the Reynolds number which can be, then, evaluated as 
$Re = U_{\text{cap}}h_0/\nu$ \cite{RevModPhys.69.931}, whereas for $\Gamma > 1$, $Re = v_{\theta} h_0/\nu$. 
In any case, the Reynolds number stays always below $\approx 0.2$.
We have added the following comment at the end of the {\it Method} section:\\

\textcolor{red}{The Reynolds number, $Re$, is of the order of $Re \sim 10^{-2}$ on the 
static substrate and never exceeds the value $\approx 0.2$ in the time dependent case.}

}

\item[ \textbf{\underline{Comment 4.}}]
{ 
In addition, there is a lot of old literature that relates the dewetting to the disjoining pressure, and shows that the nucleation of a hole in a wetting film is much more difficult than the nucleation of such a wetting film; there are experiments on this by Rutledge and Taborek, theory by Schick and Taborek and Bausch, Blossey and Indekeu.
}

\item[ \textbf{{Answer}}]
{
We are well aware of the literature quoted by the Referee, but these are works 
about {\it prewetting}, namely the phenomenon of {\it nucleation} (condensation)
of films/droplets on solid surfaces confining a critical liquid-vapour mixture
{\it as a function of temperature}. So, we are a bit puzzled about what such literature may have to do with the {\it dewetting} of a film on a chemically patterned substrate. Here, in fact, we are modelling a thin liquid film which is 
thermodaynamically stable against evaporation (i.e. one can consider that 
we are at a - fixed - temperature below the critical one).
}

\item[ \textbf{\underline{Comment 5.}}]
{ 
The explanations also lack a discussion on time, length, and energy scales. 
I have no clue what the values mean ``the numerical values, in LB units, are set to $\gamma= 0.01$ and $\mu = 1/6$''?. 
}

\item[ \textbf{{Answer}}]
{
We agree that giving the values of surface tension and dynamic viscosity 
in lattice Boltzmann units was not particularly insightful and could, on 
the contrary, be misleading, so we have removed the reference to them.
\JH{But shouldn't we list them somewhere to make sure the paper is reproducible?}
Concerning length and time scales, as we have discussed in Comment 2, 
they are made dimensionless 
with the mean film height and the characteristic time $t_0$. These two quantities 
have a clear physical meaning and can, therefore, be used to compare with 
experimental setups. In particular, the expression of $t_0$, through its dependence 
on $q_0$ (the wavenumber of the most unstable mode) contains the details 
of the disjoining pressure, such as the Hamaker constant $A_H$. For instance, let us consider the problem studied in \cite{becker2003complex}, the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon 
dioxide substrate. The Hamaker constant for this system is 
$A_H  = 2.2 \times 10^{-20} \, \text{J}$ and the surface tension is 
$\gamma = 0.03 \, \text{N}/\text{m}$, such that, 
from equation (3) of \cite{becker2003complex} and 
from the definition of $q_0^2 = \frac{1}{2\gamma} \Pi^{\prime}(h_0)$, 
we get (neglecting the short range part of the interface potential)
$$
q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}.
$$
The characteristic time of the growth of the instability, then, reads \cite{PhysRevLett.99.114503} (the dynamic viscosity of polystyrene is 
$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$)
$$
t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}
$$
and it can be seen that indeed film rupture occurs within time 
scales of the order of $t_0$ ($\approx 2 t_0$) \cite{becker2003complex}.
We have reported these comments on comparison with actual experimental 
values in the new version of the paper.
}

\item[ \textbf{\underline{Comment 6.}}]
{ 
What is the disjoining pressure due to? 
Van der Waals forces? 
How thick is the precursor film, and why should the disjoining pressure be zero there?
Is there any comparison with experiments possible/feasible?
}

\item[ \textbf{{Answer}}]
{
As we have already discussed in the first answer (and as one can find 
in several reviews and undergraduate textbooks, but see, e.g., Craster 
and Matar \cite{craster2009dynamics}), the disjoining pressure encode 
any kind of interactions among the fluid molecules and the substrate, so 
also those of Van der Waals type. The precursor film is an equilibrium value for the thickness that emerges when the interface potential has also a short 
range repulsive term. The disjoining pressure is zero there by definition of equilibrium (or, equivalently, because that is a minimum of the interface potential). Concerning comparisons of a thin-film equation approach with 
experiments, there are tons of examples in the literature, such as 
\cite{becker2003complex,fetzer2005new,PhysRevLett.99.114503} to name but a few (see also \cite{craster2009dynamics,RevModPhys.81.739} for reviews). 
}

\end{itemize}

\bibliographystyle{abbrv}
\bibliography{Ref}

\end{document}

### 2022-10-03 main_AS.tex
Update

%%
%\listfiles
%\documentclass[%
% reprint,%
%secnumarabic,%
% amssymb, amsmath,%
% aip,cha,%
%groupedaddress,%
%frontmatterverbose,
%]{revtex4-1}
\documentclass[twocolumn,amsmath,amssymb,showpacs,pre,nofootinbib,superscriptaddress]{revtex4-1} %,preprint,jcp

\bibstyle{apsrev}
%\usepackage{docs}%
\usepackage{bm}%
\usepackage{graphicx}
%\usepackage{subcaption}
\usepackage{tikz}
\usepackage{color}
\usepackage{xcolor}
\usepackage{physics}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}%
%\nofiles
\expandafter\ifx\csname package@font\endcsname\relax\else
 \expandafter\expandafter
 \expandafter\usepackage
 \expandafter\expandafter
 \expandafter{\csname package@font\endcsname}%
\fi
\hyphenation{title}
\newcommand{\JH}[1]{\textcolor{blue}{ JH: #1}}

@@

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

%\title{Switchable substrates and thin film morphologies}
%\title{Time varying contact angles and thin film morphologies}
\title{Morphologies of thin liquid film dewetting on time-dependent patterned substrates}

\author{S. Zitz}
\email{s.zitz@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
  \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\author{A. Scagliarini}%
\email{andrea.scagliarini@cnr.it}
 \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), \\
Consiglio Nazionale delle Ricerche (CNR),\\
Via dei Taurini 19, 00185 Rome, Italy}%
\affiliation{INFN, sezione Roma ``Tor Vergata'', via della Ricerca Scientifica 1, 00133 Rome, Italy}
\author{J. Harting}
\email{j.harting@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,\\
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
 \affiliation{Department of Chemical and Biological Engineering and Department of Physics, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\date{\today}

\begin{abstract}
We study the dynamics of an unstable thin film deposited on a ``switchable`` substrate.
The substrate is itself subject to a time varying function that affects the contact angle $\theta(\mathbf{x},t)$.
We show that in the case of vanishing time dependence the patterning drives the dewetting morphology and that all fluid is drained into regions of small $\theta(\mathbf{x})$ and thus form droplets.
Interestingly enough with a time varying component to the pattern we are able to stop the droplet fragmentation and create pronounced ligament or rivulet morphologies.
We show that both the numbers of semistable rivulets as well as number of stable droplets can be controlled with the pattern wavelength ($\lambda$).
Finally we explain why the ligaments appear with a simple balance between capillary retraction and dynamic wetting.   
\end{abstract}

@@

\newcommand{\ts}{\textsuperscript}
%\tableofcontents

\section{Introduction}\label{sec:intro}
Wet surfaces and droplets are part of our every-day experience and of several industrial processes,
such as coating, tribology, painting and printing,
to name but a few~\cite{gross1980fluid,szeri2010fluid,DERYCK1998278,doi:10.1146/annurev.fluid.31.1.347,DASILVASOBRINHO19991204,singh2010inkjet,jo2009evaluation}. 
%Being it the droplets condensed in the higher atmosphere and return to the ground as rain or the oil film in a pan during cooking. 
%While this two examples are illustrative there are many industrial applications that build on thin film flows. 
%Among them is the wetting of surfaces to reduce wear or put simply, lubrication~\cite{ReynoldsLubr, gross1980fluid, szeri2010fluid}. 
%Coating surfaces to change their physio-chemical properties is another widely used application~\cite{DERYCK1998278, doi:10.1146/annurev.fluid.31.1.347, DASILVASOBRINHO19991204}.
%On the other hand the need to control droplets is at the heart of many printing processes~\cite{singh2010inkjet, jo2009evaluation}.
Moreover, the continuously growing technological interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for
printable electronics~\cite{Kim_2005, Luechinger_2008}, whose efficiency relies crucially on a precise control of material deposition upon dewetting of
liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e. it is not inert but responds
dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{doi:10.1126/science.288.5471.1624, doi:10.1246/bcsj.20180076, xin2010reversibly}.
%More recently lab-on-a-chip devices have acquired much attention, to work they require precise droplet or
%film flow on small scales~\cite{stone2004engineering, darhuber2010planar}.
%One possibility for precise control is the local wettability.
%Light switchable substrates~\cite{WANG200718}, therefore substrates that change their wettability based on light illumination, open up a new% way to program this devices.

Spatial variations of the contact angle are long known and used in previous studies~\cite{PhysRevE.91.023013, Sommers_2006}, time dependency of the contact angle on the other hand is not as well studied~\cite{suman2006dynamics}. 
Switchable substrates are in fact a realization of a time dependent contact angle. 
Using both spatial as well as time dependent variations of the local contact angle leads to interesting phenomena, for example continuously sliding droplets as shown by Grawitter and Stark~\cite{D0SM02082F}.  
Our approach builds on the same idea but our method, see ref~\cite{PhysRevE.100.033313}, allows us to go beyond a single droplet and to simulate the full dewetting transition starting with a thin film and ending with droplets.
We use this to show that a time dependent component of the contact angle alter the dewetting morphology quite substantially.
In the limit of a small time dependency or slow evolution of the contact angle field we qualitatively reproduce the results from Grawitter and Stark, such sliding droplets.
Speeding up the evolution of the contact angle field introduces a morphological transition, which to the understanding of the authors have not been looked at so far.
Therefore the purpose of this manuscript is to increase the understanding of the dewetting process.
Highlighting the fact that a temporal and spatial variation of the contact angle leads to novel morphological structures.
Thus our finding offer a simple way to understand how time periodic switchable substrates could generate novel dewetting morphologies.

The manuscript is outlined as follows:
In section~\ref{sec:method} we discuss the numerical method we use to generate our results and introduce relevant scales.
We then present our results in section~\ref{sec:results}, where we discuss the impact of the time dependency on the contact angle and show that both stability as well as dewetting morphology is affected.
A short summary of our finding and a closing conclusion with outlook is given in section~\ref{sec:sum_conclu}.

\section{Numerical model and simulation details}\label{sec:method}
In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{ReynoldsLubr,RevModPhys.69.931,RevModPhys.81.1131} 

@@

by means of a recently developed method, based on a lattice Boltzmann
scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801}.
Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface;
the mobility function $M_{\delta}(h)$ depends on the velocity boundary condition at the substrate, parameterized by an effective slip length $\delta$, and reads

@@

  M_{\delta}(h) = \frac{2h^3 + 6\delta h^2 + 3\delta^2h}{6\mu},

@@

(which, for $\delta \rightarrow 0$, reduces to the no-slip form $h^3/(3\mu)$),
where $\mu$ is the fluid dynamic viscosity.
The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$. The disjoining pressure
$\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential; as such, it contains the information on the
liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle, $\theta$ \cite{RevModPhys.81.739, SCHWARTZ1998173}. 
The expression adopted for $\Pi$ is:

@@

  \Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))\left[\left(\frac{h_{\ast}}{h}\right)^3 - \left(\frac{h_{\ast}}{h}\right)^2\right],

@@

where $h_{\ast}$, the height at which the disjoining pressure vanishes, sets the precursor layer thickness.
The time variability of the patterned substrate, then, enters the model precisely through the
disjoining pressure, by making the contact angle space and time dependent, $\theta = \theta(\mathbf{x},t)$.
In particular, the following sinusoidal form is employed:

@@

    \theta(\mathbf{x},t) = \theta_0 + \delta\theta\left[\sin\left(\frac{2\pi (x+v_{\theta x}t)}{\lambda}\right)\sin\left(\frac{2\pi (y+v_{\theta y}t)}{\lambda}\right)\right], 

@@

with $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$; the pattern evolves in time as a plane wave.
We choose to fix the velocity direction to one diagonal, i.e.
$\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},1/\sqrt{2})$.
Notice that, since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one time step the wave would travel
a distance much smaller than a lattice spacing), the time update must be interpreted in an integer part sense, that is
the pattern is shifted by one $\Delta x$ every $1/v_{\theta x}$ time steps (and equivalently in the $y$-direction).
%The second pattern is a triangle wave which accounts for a different kind of pattern gradient.
%To the best of our knowledge contact angles can not be tuned precise enough to match a sinusoidal wave on %the scales we are interested in.
%Therefore we assume that a triangle wave with a linear gradient is more likely accessable in experiments, %thus $\theta_2$ is given as
%\begin{equation}\label{eq:triangle_wave}
%\begin{split}
%    \theta_2(x,y) &= \theta_0 + \\ &\frac{2\delta\theta}{\pi}\left[\sin^{-1}\left(\sin\left(\frac{2\pi x}{%\lambda}\right)\right)\sin^{-1}\left(\sin\left(\frac{2\pi y}{\lambda}\right)\right)\right].
%    \end{split}

>>>>

\documentclass[twocolumn,amsmath,amssymb,showpacs,prl,superscriptaddress]{revtex4-1} %,preprint,jcp,nofootinbib

\bibstyle{apsrev}
\usepackage{bm}%
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{color}
\usepackage{xcolor}
\usepackage{physics}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}%
\expandafter\ifx\csname package@font\endcsname\relax\else
 \expandafter\expandafter
 \expandafter\usepackage
 \expandafter\expandafter
 \expandafter{\csname package@font\endcsname}%
\fi
\hyphenation{title}
\newcommand{\JH}[1]{\textcolor{blue}{ JH: #1}}

@@

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}
\definecolor{jlblue}{rgb}{0.0,0.6056031611752245,0.9786801175696073}
\definecolor{jlorange}{rgb}{0.8888735002725198,0.43564919034818994,0.2781229361419438}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

\title{Controlling the dewetting morphologies of thin liquid films by switchable substrates}

\author{S. Zitz}
\email{zitz@ruc.dk}
 %Changed order: Work done while Stefan worked in Nuremberg
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
  \affiliation{Department of Chemical and Biological Engineering, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
  \affiliation{IMFUFA, Department of Science and Environment,\\ 
Roskilde University, Postbox 260, DK-4000 Roskilde, Denmark}%
\author{A. Scagliarini}%
\email{andrea.scagliarini@cnr.it}
 \affiliation{Institute for Applied Mathematics "M. Picone" (IAC), 
Consiglio Nazionale delle Ricerche (CNR),\\
Via dei Taurini 19, 00185 Rome, Italy}%
\affiliation{INFN, sezione Roma ``Tor Vergata'', via della Ricerca Scientifica 1, 00133 Rome, Italy}
\author{J. Harting}
\email{j.harting@fz-juelich.de}
 \affiliation{Helmholtz Institute Erlangen-N\"urnberg for Renewable Energy,\\
  Forschungszentrum J\"ulich,
  F\"urther Strasse 248, 90429 N\"urnberg, Germany}%
 \affiliation{Department of Chemical and Biological Engineering and Department of Physics, Friedrich-Alexander-Universit\"at Erlangen-N\"urnberg, F\"{u}rther Stra{\ss}e 248, 90429 N\"{u}rnberg, Germany}
\date{\today}

\begin{abstract}
\noindent Switchable and adaptive substrates emerged as valuable tools for the control of wetting and actuation of droplet motion. Here we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modelled with a space and time varying contact angle.
With a static pattern, all the fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a state consisting of metastable rivulets is observed. 
A criterion discriminating whether rivulets can be observed or not is identified in terms of a single dimensionless parameter.
Finally, we show and explain theoretically how the film rupture times, droplet shape and rivulet life time depend on the pattern wavelength and speed.  
\end{abstract}

@@

\newcommand{\ts}{\textsuperscript}

\noindent {\it Introduction}. Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{gross1980fluid,szeri2010fluid,DERYCK1998278,doi:10.1146/annurev.fluid.31.1.347,DASILVASOBRINHO19991204,singh2010inkjet,jo2009evaluation,Wijshoff2010}. 
Moreover, the continuously growing technological interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for printable electronics~\cite{Kim_2005, Luechinger_2008} and printable photovoltaics~\cite{Brabec2001,RH20}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e. it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{ButtEtAl_Langmuir2018,GuoGuo_RSCAdv2016}. 
Several realizations of switchable and adaptive substrates have been proposed~\cite{LiuChem2005, XiCSR2010}, involving smart materials such as polymer brushes~\cite{CohenStuartEtAl_NatMat2010,ayres2007stimuli,YongEtAl_Mat2018}, thermal-responsive hydrogels~\cite{ChenEtAl_SM2010}, light-responsive molecules and microstructures~\cite{IchimuraEtAl_Science2000,DelormeEtAl_Langmuir2005,OscuratoEtAl_AAMI2017} or processes such as electrowetting~\cite{MugeleEtAl_JPCM2005}.
Probably the simplest, yet non-trivial, modelling of a switchable substrate can be realized by a space and time dependent wettability pattern~\cite{GrawitterStark1}. 
While a consistent body of theoretical/computational work has been devoted to processes on static heterogeneous substrates, though, the time dependent case is still almost unexplored, with few relevant exceptions focusing on single droplet spreading and sliding~\cite{GrawitterStark1,GrawitterStark2,ThieleHartmann} or limited to an analysis of the linear regime~\cite{suman2006dynamics}.

In this Letter, we study, by means of numerical simulations, the full dewetting dynamics of a thin liquid film deposited over a substrate with a time varying wettability pattern, from rupture to the long time morphology. 
We identify two regimes where the rupture times grow with the pattern wavelength either linearly (on the static pattern) or attain a constant value (in the time dependent case), for short wavelengths, and approach a quadratic law as the wavelength increases. These observations are then explained theoretically.
We show that, by tuning the rate of change (the ``speed'') of the underlying pattern, it is possible to control, to some extent, the dewetting morphology. In particular, for large enough pattern speeds, we detect a metastable state, where the film retracts into metastable rivulets, eventually breaking up into multiple droplets.
We introduce a control parameter that discriminates whether rivulets or just droplets (as in the static situation)
can be observed and we propose a phenomenological argument that justifies the logarithmic dependence of the rivulets life time on the pattern speed.\\
\\
\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{ReynoldsLubr,RevModPhys.69.931,RevModPhys.81.1131} 

@@

by means of a recently developed method, based on a lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801}.
Eq.~(\ref{eq:thinfilm}) describes, in a lubrication approximation spirit, the evolution of the height field (film thickness) $h(\mathbf{x},t)$, denoting the location of the liquid/air interface. The mobility function 

@@

  M_{\delta}(h) = \frac{2h^3 + 6\,\delta\, h^2 + 3\delta^2h}{6\mu},

@@

depends on the velocity boundary condition at the substrate, parameterized by an effective slip length $\delta$
(for $\delta \rightarrow 0$ it reduces to the no-slip form $h^3/(3\mu)$). Here, $\mu$ is the fluid dynamic viscosity.
The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$, 
\textcolor{red}{where $\gamma$ is the surface tension}.
The disjoining pressure $\Pi$ can be seen as (minus) the derivative, with respect to the film thickness, of an effective interfacial potential.
As such, it contains the information on the liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle $\theta$~\cite{RevModPhys.81.739, SCHWARTZ1998173}. 
The expression adopted for $\Pi$ is

@@

\textcolor{red}{\Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))
  f\left(\frac{h}{h_{\ast}}\right),}

@@

\textcolor{red}{where $f(\xi)=\xi^{-3} - \xi^{-2}$} and $h_{\ast}$ is the height at which the disjoining pressure vanishes and sets the precursor layer thickness.
The time variation of the patterned substrate then enters the model precisely through the disjoining pressure, by making the contact angle space and time dependent, i.e. $\theta = \theta(\mathbf{x},t)$.
In particular, we employ the sinusoidal form

@@

   \!\! \theta(\mathbf{x},t) = \theta_0 + \delta\theta\left[\sin\left(q_{\theta} (x+v_{\theta x}t)\right)\sin\left(q_{\theta}(y+v_{\theta y}t)\right)\right],\! 

@@

where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave. 
We fix the velocity direction to one diagonal, namely $\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},-1/\sqrt{2})$ (we will return later on the importance of this choice), and we set $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$~\footnote{Notice that, since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one LB time step $\Delta t$ the wave would travel a distance much smaller than a lattice spacing $\Delta x$),
the time update needs to be interpreted in an integer part sense, that is the pattern is shifted by one $\Delta x$ every $1/v_{\theta x}$ time steps (and equivalently in the $y$-direction).}.
\textcolor{red}{Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film \cite{Mecke_2005}. On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads 
